Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2020-03-10 ICML2019/main.tex
update the theory prob of laplace mecahism and exp mechanism
update v 2020
prelimilary

A more natural approach, which we propose here, is to also directly measure sensitivity with respect
to a metric on the space of inferred probability distributions. Our goal is to identify cases where jointly measuring
accuracy and sensitivity in terms of a single metric on probability distributions can yield improved accuracy. 
{\color{red}{MB: Is this accurate? Would we ever want to use different metrics for accuracy and sensitivity?}}

@@

%  support bayesian inference
% by means of a language primitive, often called $\bysinfer$, which
% supports the (precise or approximate) parametric computation of the
% posterior of a given model. In the rest of the paper, we will use the following
% notation as a shorthand for the inference process:
% \[
%   \bysinfer(\Pr(\dataobs | \theta), \Pr(\theta), \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
% \]
\noindent \textbf{Dirichlet-multinomial model.}
In this work we will focus on the Dirichlet-Multinomial model for
categorical data and its restriction to binary data: the Beta-Binomial
model. 
That is, we will consider the situation where 
the underlying categorical data is drawn from a multinomial
distribution with parameter the vector $\vtheta\in [0,1]^{k}$. The prior distribution over $\vtheta$
is given by a Dirichlet distribution, $\dirichlet{\valpha}$, for,
$\valpha\in(\mathbb{R}^{+})^{k}$ and $k\in\mathbb{N}$, with p.d.f:
\[
\Pr(\vtheta)=\frac{1}{\mbetaf(\valpha)}\cdot\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generalized beta function.
The data $\dataobs$ is a sequence of $n\in\mathbb{N}$ values
coming from a universe $\datauni$, such that $|\datauni|=k$.
The likelihood function is
$
\Pr(\dataobs|\vtheta)=\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
$
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\mathbf{1}_{\iverson{x_j=a_i}}$.
Denoting by $\Delta\valpha$ the vector $(\Delta\alpha_1,\dots
\Delta\alpha_k)$ the posterior distribution over $\vtheta$ is
$\dirichlet{\valpha + \Delta \valpha}$, where $+$ denotes the componentwise sum of vectors.
In the following we will use the notation $\bysinfer(\valpha,\vtheta,
\dataobs)$ to denote the posterior of this process.
%i.e., $\bysdir(n,\vtheta, \valpha, \dataobs)=\dirichlet{\valpha + \Delta \valpha}$.
%
% ($\thicksim \binomial{n}{\theta}$), where $\theta$ represents
% the parameter --informally called \emph{bias}-- of a Bernoulli
% distributed random variable and $n$ is the number of draws. The
% prior distribution over $\theta\in [0,1]$ is going to be a beta
% distribution, $\betad(\alpha, \beta)$, with parameters
% $\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
% \[
%   \Pr(\theta)= \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
% \]
% where $\betaf(\cdot,\cdot)$ is the beta function.
% The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
% \[
%   \Pr(\dataobs | \theta)= \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
% \]
% where $\Delta \alpha = \sum_{i=1}^{n}x_i$.
% The posterior distribution over $\theta$ can be easily proven to be equal to:\[\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)\]
% From now on we will denote by $\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$ the previous
% posterior distribution. This notation will prove itself handy when its arguments will change types in the next example and
% in section \ref{} when the private mechanisms will be described.

@@

\begin{thm}
  \label{thm_gamma_smooth}
  For any $\dataobs, \dataobs'\in\mathcal{X}^n$, and $\gamma \in \mathbb{R}^{+}$
  if  $\adj{\dataobs}{\dataobs'}$ then $\frac{1}{S(\dataobs)} - \frac{1}{S(\dataobs')} \leq \gamma$.
\end{thm}
again, the interested reader will find the proof in Appendix.
We are now ready to describe the mechanism $\hexpmech$ in Algorithm \ref{mech:sexpmech}. $\hexpmech$ outputs a candidate $r \in \candidateset{\valpha}$ with probability
\[
 \frac
  {\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r)}{4\cdot S(\dataobs)})}
{\Sigma_{r' \in \candidateset{\valpha}}\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r')}{4\cdot S(\dataobs)})}
\]
where $S(\dataobs)$ is the 1-smooth sensitivity of  the function $\hellinger(\bysinfer{\dataobs},\cdot)$. {\color{red}{why do we choose 1???}}
  \begin{algorithm}
  \caption{$\hexpmech$}
  \label{mech:sexpmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, prior: $\dirichlet{\valpha}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'}= \bys{\dataobs}{\valpha}$.   
  \STATE \quad {\bf set} $z=r$ with probability $\frac{\exp(\frac{\epsilon\cdot \ux{r}}{4\cdot S(\dataobs)})}{\Sigma_{r' \in \candidateset{\valpha}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{4\cdot S(\dataobs)})}$
\STATE {\bf return} $z$
  \end{algorithmic}
\end{algorithm}

@@

\subsection{Accuracy Bound for $\hexpmech$}
\label{subsec_accuracy_smoo}
\begin{lem}
\label{lem_acc_hexp}
  In Beta-binomial model,
  we have following accuracy bound for this mechanism:
%  \begin{align*}
%  \label{eq_acc}
%  & \hexpmechPr{\dataobs}{\hellinger(\bysinfer{\dataobs}, z) = c  }\\ 
%  = & \frac{\exp{(\frac{- \epsilon c}{4 S(\dataobs)})}}{\nomalizer{\dataobs}},
%  \end{align*}
%
%  $c = \hellinger(\betad(\alpha, \beta), \betad(\alpha + t, \beta - t)$
  \begin{align*}
  \lefteqn{ \hexpmechPr{\dataobs}{\hellinger(\bysinfer{\dataobs}, z) = c(t)  } }\\ 
  &  = \frac{\exp{(\frac{- \epsilon c}{4 S(\dataobs)})}}{\nomalizer{\dataobs}}
%  & \text{ where } c(t) = \hellinger(\betad(\alpha, \beta), \betad(\alpha + t, \beta - t).
  \end{align*}
\end{lem}

>>>>

A more natural approach, which we propose here, is to also directly measure sensitivity with respect
to a metric on the space of inferred probability distributions. Our goal is to identify cases where jointly measuring
accuracy and sensitivity in terms of a single metric on probability distributions can yield improved accuracy. 
% {\color{red}{MB: Is this accurate? Would we ever want to use different metrics for accuracy and sensitivity?}}

@@

%  support bayesian inference
% by means of a language primitive, often called $\bysinfer$, which
% supports the (precise or approximate) parametric computation of the
% posterior of a given model. In the rest of the paper, we will use the following
% notation as a shorthand for the inference process:
% \[
%   \bysinfer(\Pr(\dataobs | \theta), \Pr(\theta), \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
% \]
\noindent \textbf{Dirichlet-multinomial model.}
In this work we will focus on the Dirichlet-Multinomial model for
categorical data and its restriction to binary data: the Beta-Binomial
model. 
That is, we will consider the situation where 
the underlying categorical data is drawn from a multinomial
distribution with parameter the vector $\vtheta\in [0,1]^{k}$. The prior distribution over $\vtheta$
is given by a Dirichlet distribution, $\dirichlet{\valpha}$, for,
$\valpha\in(\mathbb{R}^{+})^{k}$ and $k\in\mathbb{N}$, with p.d.f:
\[
\Pr(\vtheta)=\frac{1}{\mbetaf(\valpha)}\cdot\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generalized beta function.
The data $\dataobs$ is a sequence of $n\in\mathbb{N}$ values
coming from a universe $\datauni$, such that $|\datauni|=k$.
The likelihood function is
$
\Pr(\dataobs|\vtheta)=\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
$
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\mathbf{1}_{\iverson{x_j=a_i}}$.
Denoting by $\Delta\valpha$ the vector $(\Delta\alpha_1,\dots
\Delta\alpha_k)$ the posterior distribution over $\vtheta$ is
$\dirichlet{\valpha + \Delta \valpha}$, where $+$ denotes the componentwise sum of vectors.
In the following we will use the notation $\bysinfer{\valpha,\vtheta,
\dataobs}$ to denote the posterior of this process.
%i.e., $\bysdir(n,\vtheta, \valpha, \dataobs)=\dirichlet{\valpha + \Delta \valpha}$.
%
% ($\thicksim \binomial{n}{\theta}$), where $\theta$ represents
% the parameter --informally called \emph{bias}-- of a Bernoulli
% distributed random variable and $n$ is the number of draws. The
% prior distribution over $\theta\in [0,1]$ is going to be a beta
% distribution, $\betad(\alpha, \beta)$, with parameters
% $\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
% \[
%   \Pr(\theta)= \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
% \]
% where $\betaf(\cdot,\cdot)$ is the beta function.
% The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
% \[
%   \Pr(\dataobs | \theta)= \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
% \]
% where $\Delta \alpha = \sum_{i=1}^{n}x_i$.
% The posterior distribution over $\theta$ can be easily proven to be equal to:\[\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)\]
% From now on we will denote by $\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$ the previous
% posterior distribution. This notation will prove itself handy when its arguments will change types in the next example and
% in section \ref{} when the private mechanisms will be described.

@@

\begin{thm}
  \label{thm_gamma_smooth}
  For any $\dataobs, \dataobs'\in\mathcal{X}^n$, and $\gamma \in \mathbb{R}^{+}$
  if  $\adj{\dataobs}{\dataobs'}$ then $\frac{1}{S(\dataobs)} - \frac{1}{S(\dataobs')} \leq \gamma$.
\end{thm}
again, the interested reader will find the proof in Appendix.
We are now ready to describe the mechanism $\hexpmech$ in Algorithm \ref{mech:sexpmech}. $\hexpmech$ outputs a candidate $r \in \candidateset{\valpha}$ with probability
\[
 \frac
  {\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r)}{4\cdot S(\dataobs)})}
{\Sigma_{r' \in \candidateset{\valpha}}\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r')}{4\cdot S(\dataobs)})}
\]
where $S(\dataobs)$ is the 1-smooth sensitivity of  the function $\hellinger(\bysinfer{\dataobs},\cdot)$. 
% {\color{red}{why do we choose 1???}}
  \begin{algorithm}
  \caption{$\hexpmech$}
  \label{mech:sexpmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, prior: $\dirichlet{\valpha}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'}= \bys{\dataobs}{\valpha}$.   
  \STATE \quad {\bf set} $z=r$ with probability $\frac{\exp(\frac{\epsilon\cdot \ux{r}}{4\cdot S(\dataobs)})}{\Sigma_{r' \in \candidateset{\valpha}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{4\cdot S(\dataobs)})}$
\STATE {\bf return} $z$
  \end{algorithmic}
\end{algorithm}

@@

\subsection{Accuracy Bound for $\hexpmech$}
\label{subsec_accuracy_smoo}
\begin{lem}
\label{lem_acc_hexp}
  In Beta-binomial model,
  we have following accuracy bound for this mechanism:
%  \begin{align*}
%  \label{eq_acc}
%  & \hexpmechPr{\dataobs}{\hellinger(\bysinfer{\dataobs}, z) = c  }\\ 
%  = & \frac{\exp{(\frac{- \epsilon c}{4 S(\dataobs)})}}{\nomalizer{\dataobs}},
%  \end{align*}
%
%  $c = \hellinger(\betad(\alpha, \beta), \betad(\alpha + t, \beta - t)$
  \begin{align*}
  \lefteqn{ \hexpmechPr{\dataobs}{\hellinger(\bysinfer{\dataobs}, z) = c(t)  } }\\ 
  &  = \frac{\exp{(\frac{- \epsilon c(t)}{4 S(\dataobs)})}}{\nomalizer{\dataobs}}
%  & \text{ where } c(t) = \hellinger(\betad(\alpha, \beta), \betad(\alpha + t, \beta - t).
  \end{align*}
\end{lem}
